### 2021-01-26 background.tex
update reference and study setting begin

The idea of mine automatically software resources or components, to inferring behavior is not new, and has been discussed before. For instance, Whaley et al. \cite{whaley2002automatic} combine dynamic and static analysis to API mining and so extract interface from software components. Ammons et al \cite{ammons2002mining} propose a machine learning
approach, called specification mining, to discovering some temporal and data-dependence relationships that a program follows when interacting with an API or abstract data type.

Regarding test generating tools used to mining Sandboxes, Jamrozik et al \cite{jamrozik2016droidmate} proposed DroidMate, a test generation tools that implements a pseudo-random GUI exploration strategy, and was the first approach to leverage test generation to extract sandbox rules from apps. Li e tal. proposed in \cite{li2017droidbot}, DroidBot, a test generator tool that explores sensitive resources access by apps, following a model-based exploration strategy. In this paper, the authors presented comparison between DroidBot and Monkey \cite{Monkey} in a proof of concept example of using DroidBot in malware analysis. From the same authors, another test generator tool for Android, described as Humanoid \cite{humanoid-paper}, is a Droidbot evolution and presents a proposal that can generate humans like tests inputs, using deep learning.

L Bao et al. \cite{bao2018mining} had provided an important comparative study between five test generation tools in finding malware using mining sandboxes techniques: DroidMate \cite{jamrozik2016droidmate}, Monkey \cite{Monkey}, GUIRipper \cite{amalfitano2012using}, Puma \cite{hao2014puma}, and Droidbot \cite{li2017droidbot}. This study indicates that these tools was efficient in identifying at most 70\% of the malware in a specific dataset and also reports that, even after combining all test generator tools, it was possible to detect 80\% of malicious behavior. However, this study did not focus on the possible interference of static analysis in the final result, since this study used a instrumentation tool \cite{cai2017droidfax}, that also performs a static analysis of the apps.

>>>>

The idea of mine automatically software resources or components, to inferring behavior is not new, and has been discussed before. For instance, Whaley et al. \cite{DBLP:conf/issta/WhaleyML02} combine dynamic and static analysis to API mining and so extract interface from software components. Ammons et al \cite{DBLP:conf/popl/AmmonsBL02} propose a machine learning
approach, called specification mining, to discovering some temporal and data-dependence relationships that a program follows when interacting with an API or abstract data type.

Regarding test generating tools used to mining Sandboxes, Jamrozik et al \cite{DBLP:conf/icse/JamrozikZ16} proposed DroidMate, a test generation tools that implements a pseudo-random GUI exploration strategy, and was the first approach to leverage test generation to extract sandbox rules from apps. Li e tal. proposed in \cite{DBLP:conf/icse/LiYGC17}, DroidBot, a test generator tool that explores sensitive resources access by apps, following a model-based exploration strategy. In this paper, the authors presented comparison between DroidBot and Monkey \cite{Monkey} in a proof of concept example of using DroidBot in malware analysis. From the same authors, another test generator tool for Android, described as Humanoid \cite{DBLP:conf/kbse/LiY0C19}, is a Droidbot evolution and presents a proposal that can generate humans like tests inputs, using deep learning.

L Bao et al. \cite{DBLP:conf/wcre/BaoLL18} had provided an important comparative study between five test generation tools in finding malware using mining sandboxes techniques: DroidMate, Monkey, GUIRipper \cite{DBLP:conf/kbse/AmalfitanoFTCM12}, Puma \cite{DBLP:conf/mobisys/Hao0NHG14}, and Droidbot. This study indicates that these tools was efficient in identifying at most 70\% of the malware in a specific dataset and also reports that, even after combining all test generator tools, it was possible to detect 80\% of malicious behavior. However, this study did not focus on the possible interference of static analysis in the final result, since this study used a instrumentation tool \cite{DBLP:conf/icsm/CaiR17a}, that also performs a static analysis of the apps.

### 2021-01-26 introduction.tex
update reference and study setting begin

Almost two-thirds of the world population use mobile technologies~\cite{Comscore}, and the Android Operating System has dominated the market of smartphones, tablets, and others electronic devices. Due to this growing popularity, the number of incidents related to Android malicious software (malware) has significantly increased \cite{faruki2014android,tan2015securing}. Security issues in Android software applications (``apps'' for short) have become a relevant research topic, and many techniques have been developed to identify vulnerabilities in Android apps \cite{tan2015securing}, including the use of static analysis algorithms to identify source and sink paths involving sensitive data~\cite{arzt-pldi-2014} and to reveal the misuse of cryptographic primitives~\cite{krueger:ecoop-2018,rahaman:ccs-2019}. Using a different approach, researchers have also explored the use of dynamic analysis to mine Android sandboxes, whose goal is to protect Android users from malicious behavior~\cite{jamrozik2016mining}.

@@

Jamrozik et al.~\cite{jamrozik2016mining} argue in favor of dynamic analysis for mining sandboxes, instead of using static analysis. The authors claim that dynamic analysis exceeds static analysis on mining Android sandboxes because the analysis must often assume that additional behaviors are possible than actually would be. Otherwise, static analysis might determine that some behavior is impossible, this behavior no longer needs to be checked at runtime. Thereby, code that is interpreted, decrypted, or downloaded at run time is challenging for static analysis---even though these more \emph{dynamic} strategies are often used to introduce malicious behavior. Nonetheless, the authors do not give any empirical validation about the limitations of static analysis to mine sandboxes. In addition, it is not clear how both approaches (dynamic and static analysis) could complement each other in the process of mining Android sandboxes.

The same lack of understanding about the role of static analysis also appears in the work of Bao et al. \cite{bao2018mining}, which present an empirical study that compares different dynamic analysis tools for testing and mining Android sandboxes. Their study leverages DroidFax \cite{cai2017droidfax} to instrument 102 pairs of Android apps (each pair comprising a benign and a malign version of an App) and to collect the information needed to mine sandboxes (that is, the calls to sensitive APIs).
Although the authors report a performance of at most 70\% of dynamic analysis tools to differentiate the benign and malicious versions of the apps, the authors ignore that DroidFax statically analyzes the Android apps and records calls to sensitive APIs (besides instrumenting the apps), which might lead to an over estimation of the performance of the dynamic analysis tools for mining sandboxes and introduce a possible threat to the conclusions of that work.

Our goal is to understand how static analysis
algorithms might complement dynamic analysis approaches in the process of mining Android sandboxes. To this end, we performed two studies. In the
first, we conduct a non-exact replication of the work of Bao et al.~\cite{bao2018mining}. Our study differs from the original work in two aspects: (a)
by isolating the effect of the static analysis component of DroidFax, (b) by considering more recent dynamic analysis tools for testing Android apps,
and (c) by evaluating the performance of DroidMate~\cite{jamrozik2016mining} in its original settings. Considering the other dynamic analysis approaches, which are
general purpose test case tools for Android, DroidMate has been designed with the goal of mining Android sandboxes---eventhough it has not been tested adequatly
in the work of Bao et al~\cite{bao2018mining}. In the second study, we explore to what extent we can complement the dynamic analysis approach using
tainted analysis algorithms.

>>>>

Almost two-thirds of the world population use mobile technologies~\cite{Comscore}, and the Android Operating System has dominated the market of smartphones, tablets, and others electronic devices. Due to this growing popularity, the number of incidents related to Android malicious software (malware) has significantly increased \cite{DBLP:journals/comsur/FarukiBLGGCR15,DBLP:journals/csur/SufatrioTCT15}. Security issues in Android software applications (``apps'' for short) have become a relevant research topic, and many techniques have been developed to identify vulnerabilities in Android apps \cite{DBLP:conf/pldi/ArztRFBBKTOM14}, including the use of static analysis algorithms to identify source and sink paths involving sensitive data~\cite{DBLP:conf/pldi/ArztRFBBKTOM14} and to reveal the misuse of cryptographic primitives~\cite{krueger:ecoop-2018,rahaman:ccs-2019}. Using a different approach, researchers have also explored the use of dynamic analysis to mine Android sandboxes, whose goal is to protect Android users from malicious behavior~\cite{DBLP:conf/icse/JamrozikSZ16}.

@@

Jamrozik et al.~\cite{DBLP:conf/icse/JamrozikSZ16} argue in favor of dynamic analysis for mining sandboxes, instead of using static analysis. The authors claim that dynamic analysis exceeds static analysis on mining Android sandboxes because the analysis must often assume that additional behaviors are possible than actually would be. Otherwise, static analysis might determine that some behavior is impossible, this behavior no longer needs to be checked at runtime. Thereby, code that is interpreted, decrypted, or downloaded at run time is challenging for static analysis---even though these more \emph{dynamic} strategies are often used to introduce malicious behavior. Nonetheless, the authors do not give any empirical validation about the limitations of static analysis to mine sandboxes. In addition, it is not clear how both approaches (dynamic and static analysis) could complement each other in the process of mining Android sandboxes.

The same lack of understanding about the role of static analysis also appears in the work of Bao et al. \cite{DBLP:conf/wcre/BaoLL18}, which present an empirical study that compares different dynamic analysis tools for testing and mining Android sandboxes. Their study leverages DroidFax \cite{DBLP:conf/icsm/CaiR17a} to instrument 102 pairs of Android apps (each pair comprising a benign and a malign version of an App) and to collect the information needed to mine sandboxes (that is, the calls to sensitive APIs).
Although the authors report a performance of at most 70\% of dynamic analysis tools to differentiate the benign and malicious versions of the apps, the authors ignore that DroidFax statically analyzes the Android apps and records calls to sensitive APIs (besides instrumenting the apps), which might lead to an over estimation of the performance of the dynamic analysis tools for mining sandboxes and introduce a possible threat to the conclusions of that work.

Our goal is to understand how static analysis
algorithms might complement dynamic analysis approaches in the process of mining Android sandboxes. To this end, we performed two studies. In the
first, we conduct a non-exact replication of the work of Bao et al.~\cite{DBLP:conf/wcre/BaoLL18}. Our study differs from the original work in two aspects: (a)
by isolating the effect of the static analysis component of DroidFax, (b) by considering more recent dynamic analysis tools for testing Android apps,
and (c) by evaluating the performance of DroidMate~\cite{DBLP:conf/icse/JamrozikZ16} in its original settings. Considering the other dynamic analysis approaches, which are
general purpose test case tools for Android, DroidMate has been designed with the goal of mining Android sandboxes---eventhough it has not been tested adequatly
in the work of Bao et al~\cite{DBLP:conf/wcre/BaoLL18}. In the second study, we explore to what extent we can complement the dynamic analysis approach using
tainted analysis algorithms.

### 2021-01-26 setup.tex
update reference and study setting begin

The main goal of this research work is to replicate Bao et al. study to understand the implications of static analysis algorithms in his results. We also investigate how static analysis can improve the performance of testing tools in the task of identifying malicious behaviors.

@@

In the first experiment, we executed the benchmark considering 98 pairs of apps for 3 minutes. At this time, we used a new option of the benchmark that disables the static analysis performed by DroidFax. So, we used this experiment to answer the first research questions (RQ1). In the second experiment, we executed all apps with a fake tool called “Joke”, which does not do anything during the benchmark execution, to compute the results without the dynamic analysis influence. With these results, we can answer the last two research question (RQ2 and RQ3).

To assess the effectiveness of mining sandboxes created by each test case generation tools, we run these tools in benign apps and, based on resources sensitives access in this execution, we build their sandbox. Based on this sandbox, we investigate the capacity of detect malicious behaviors in the corresponding malign app. (Figure  \ref{fig:setup}) show all this experimental setup.

\subsection{Settings for RQ1}

\subsection{Settings for RQ2 and RQ3}

\begin{figure*}[h!]
  \includegraphics[width=0.5\textwidth]{images/setup.png}
  \label{Experiment setup}
  \caption{Experiment setup}
  \label{fig:setup}
\end{figure*}

>>>>

In this section, we present the setting of our study, whose main goal is to perform a non-exact replication of the work of Bao et al. and understand the implications of static analysis algorithms in his results. We also investigate how static analysis can improve the performance of testing tools in the task of identifying malicious behaviors.

@@

To assess the effectiveness of mining sandboxes created by each test case generation tools, we used a benchmark solution for test case generation called droidXP \cite{DBLP:conf/scam/CostaMCMVBC20}, that aims us to reproduce previous research work on mining sandboxes. (Figure \ref{fig:benchArq}) show all details of the benchmark architecture used at our experiments. The benchmark run these tools in benign apps and, based on resources sensitives access in this execution, we build their sandbox.

Thereby, to answer the research questions above, we perform two experiments. At both we employed the same dataset, 98 pair of real-world apps (B/M), shared by the AndroZoo \cite{DBLP:conf/msr/AllixBKT16} group. This time we executed each app in one of the five test generator tool, selected for our experiments, for 3 minutes, and for just one time. The droidXP benchmark also relies on DroidFax, the same tool used by Bao et al. in your study, that instruments Android apps and collects relevant information about their execution (using the test case generation tools). DroidFax collects the set of sensitive APIs during a test execution, and during a static analysis at each code apps. Based on this sensitive APIs, that will compose our sandbox, we investigate the capacity of detect malicious behaviors in the corresponding malign app by the tool under analysis, together with the static analysis of the DroidFax. (Figure \ref{fig:setup}) show this experimental setup.

\subsection{First Study: hybrid proposal}

In the first experiment, we executed the benchmark considering the droidXP default configuration, that is, using a hybrid proposal. To generate inputs to apps, we investigate five popular test case generation tools, one from industry (Monkey \cite{Monkey}) and four from academia (Droidbot \cite{DBLP:conf/icse/LiYGC17}, Humanoid \cite{DBLP:conf/kbse/LiY0C19}, DroidMate \cite{DBLP:conf/icse/JamrozikZ16} and ``Joke" tool). This last is a fake tool that we implemented, which does not do anything during the benchmark execution, and through it we can compute results without the dynamic analysis influence. So, we used this experiment to answer the first research questions (RQ1).

\subsection{Second Study: Dynamic analysis proposal}

In this second experiment we used a option of the benchmark that disables the static analysis performed by DroidFax, and with this new configuration, we performed our second study in the same set of apps from the first, at the same execution time, and at the same test case generation tools, including our ``Joke" tool. With the results, it was possible to compute what was the aid portion of the static analysis in the malware detection, since the Joke results, was computed without the dynamic analysis influence. We used this experiment to answer the research questions (RQ2) and (RQ3).

\begin{figure*}[ht]
  \includegraphics[width=0.5\textwidth]{images/setup.png}
  \label{Experiment setup}
  \caption{Experiment setup}
  \label{fig:setup}
\end{figure*}
